---
title: "Data Visualization"
author: "Michael Goulet"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prerequisites


```{r, message = FALSE}
library(tidyverse)
```


## First Steps

```{r}
mpg
```
### Creating a ggplot

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```
```{r}
ggplot(data = mpg)
```

```{r}
dim(mpg)
```

```{r}
# ?mpg
# `drv` is a factor of the drive train with 3 levels
```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = cyl, y = hwy))
```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = drv, y = class))

# Because the two variables are both categorical, the points will all overlap. We cannot get a sense of how many 4 wheel drive SUVs there are, for instance. 
```
## Aesthetic Mappings 

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```
### Exercises 
```{r}
# To set an aesthetic manually, assign it outside of the mapping
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue", shape = 10)
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
# Why isn't this blue? See above.
```

The categorical variables are `manufacturer`, `model`, `trans`, `drv`, `fl`, and `class`. The remaining are continuous.

```{r}
# mapping a continuous variable to color, size, and shape 
# NOTE: shape cannot be mapped to a continuous variable
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = cty, color = hwy, alpha = hwy, shape = fl), stroke = 0.2, size = 4)
```
```{r}
## Creating a custom shape by layering points of different properties
ggplot(data = mpg, mapping = aes(x = displ, y = cty)) +
  geom_point(colour = "black", size = 4.5) +
  geom_point(colour = "grey", size = 4) +
  geom_point(mapping = aes(shape = factor(cyl), color = cyl == 8))
```


## Common Problems 

Type carefully!

## Facets 

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ class, nrow = 2)
```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(fl ~ cyl)
```

### Exercises

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ cty)
```

#### Facets with empty panels

Certain panels are empty, because there are no observations that have, for example, `drv == 4` and `cyl == 5`. 
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl)) +
  facet_grid(cyl ~ drv)
```

### Exercises (cont.)
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid( ~ cyl)
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(cyl ~ class)
```

## Geometric Objects

```{r}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
```

```{r}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, color = drv))
```
### Exercises

Recreating graphs.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth(se = FALSE)
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth(mapping = aes(group = drv), se = FALSE)
```


```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_point() +
  geom_smooth(se = FALSE)
```


```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = drv)) +
  geom_smooth(se = FALSE)
```


```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = drv)) +
  geom_smooth(mapping = aes(linetype = drv),se = FALSE)
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(color = "white", size = 4) +
  geom_point(mapping = aes(color = drv))
```

## Statistical Transformations 



## Playing Around
```{r}

# truncate_dups <- function(s) {
#   return(gsub("\\s[12]$", "", s))
# }
# 
# ccn1 <- .GlobalEnv$ccn1_tibble
# ccn1$Treatments <- truncate_dups(ccn1$Sample.Name)
# ggplot(data = ccn1) +
#   geom_point(mapping = aes(x = Treatments, y = Rq))
```

